### 2024-11-05 arxiv/paper.tex
mac

\documentclass[10pt,a4paper,onecolumn]{article}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{authblk,etoolbox}
\usepackage{titlesec}
\usepackage{calc}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{colorlinks,breaklinks=true,
            urlcolor=[rgb]{0.0, 0.5, 1.0},
            linkcolor=[rgb]{0.0, 0.5, 1.0}}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{seqsplit}
\usepackage{xstring}

\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\usepackage{fixltx2e} % provides \textsubscript
\usepackage[
  backend=biber,
%  style=alphabetic,
%  citestyle=numeric
]{biblatex}
\bibliography{paper.bib}

% --- Splitting \texttt --------------------------------------------------

\let\textttOrig=\texttt
\def\texttt#1{\expandafter\textttOrig{\seqsplit{#1}}}
\renewcommand{\seqinsert}{\ifmmode
  \allowbreak
  \else\penalty6000\hspace{0pt plus 0.02em}\fi}

% --- Pandoc does not distinguish between links like [foo](bar) and
% --- [foo](foo) -- a simplistic Markdown model.  However, this is
% --- wrong:  in links like [foo](foo) the text is the url, and must
% --- be split correspondingly.
% --- Here we detect links \href{foo}{foo}, and also links starting
% --- with https://doi.org, and use path-like splitting (but not
% --- escaping!) with these links.
% --- Another vile thing pandoc does is the different escaping of
% --- foo and bar.  This may confound our detection.
% --- This problem we do not try to solve at present, with the exception
% --- of doi-like urls, which we detect correctly.

\makeatletter
\let\href@Orig=\href
\def\href@Urllike#1#2{\href@Orig{#1}{\begingroup
    \def\Url@String{#2}\Url@FormatString
    \endgroup}}
\def\href@Notdoi#1#2{\def\tempa{#1}\def\tempb{#2}%
  \ifx\tempa\tempb\relax\href@Urllike{#1}{#2}\else
  \href@Orig{#1}{#2}\fi}
\def\href#1#2{%
  \IfBeginWith{#1}{https://doi.org}%
  {\href@Urllike{#1}{#2}}{\href@Notdoi{#1}{#2}}}
\makeatother

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

% --- Page layout -------------------------------------------------------------
\usepackage[top=3.5cm, bottom=3cm, right=1.5cm, left=1.0cm,
            headheight=2.2cm, reversemp, includemp, marginparwidth=4.5cm]{geometry}

% --- Default font ------------------------------------------------------------
\renewcommand\familydefault{\sfdefault}

% --- Style -------------------------------------------------------------------
\renewcommand{\bibfont}{\small \sffamily}
\renewcommand{\captionfont}{\small\sffamily}
\renewcommand{\captionlabelfont}{\bfseries}

% --- Section/SubSection/SubSubSection ----------------------------------------
\titleformat{\section}
  {\normalfont\sffamily\Large\bfseries}
  {}{0pt}{}
\titleformat{\subsection}
  {\normalfont\sffamily\large\bfseries}
  {}{0pt}{}
\titleformat{\subsubsection}
  {\normalfont\sffamily\bfseries}
  {}{0pt}{}
\titleformat*{\paragraph}
  {\sffamily\normalsize}

% --- Header / Footer ---------------------------------------------------------
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
%\renewcommand{\headrulewidth}{0.50pt}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[L]{\hspace{-0.75cm}\includegraphics[width=5.5cm]{logo.png}}
\fancyhead[C]{}
\fancyhead[R]{}
\renewcommand{\footrulewidth}{0.25pt}

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam , Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannuksela, Neha Singh, and David Keitel, (2024). \texttt{gwsnr}:
A Python Package for Efficient SNR Calculation of Gravitational
Waves. \textit{PendingJournal of Open Source Software}, Pending(Pending), Pending. \url{https://doi.org/10.xxxxxx/draft}}}}

\fancyfoot[R]{\sffamily \thepage}
\makeatletter
\let\ps@plain\ps@fancy
\fancyheadoffset[L]{4.5cm}
\fancyfootoffset[L]{4.5cm}

% --- Macros ---------

\definecolor{linky}{rgb}{0.0, 0.5, 1.0}

\newtcolorbox{repobox}
   {colback=red, colframe=red!75!black,
     boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=3pt, bottom=3pt}

\newcommand{\ExternalLink}{%
   \tikz[x=1.2ex, y=1.2ex, baseline=-0.05ex]{%
       \begin{scope}[x=1ex, y=1ex]
           \clip (-0.1,-0.1)
               --++ (-0, 1.2)
               --++ (0.6, 0)
               --++ (0, -0.6)
               --++ (0.6, 0)
               --++ (0, -1);
           \path[draw,
               line width = 0.5,
               rounded corners=0.5]
               (0,0) rectangle (1,1);
       \end{scope}
       \path[draw, line width = 0.5] (0.5, 0.5)
           -- (1, 1);
       \path[draw, line width = 0.5] (0.6, 1)
           -- (1, 1) -- (1, 0.6);
       }
   }

% --- Title / Authors ---------------------------------------------------------
% patch \maketitle so that it doesn't center
\patchcmd{\@maketitle}{center}{flushleft}{}{}
\patchcmd{\@maketitle}{center}{flushleft}{}{}
% patch \maketitle so that the font size for the title is normal
\patchcmd{\@maketitle}{\LARGE}{\LARGE\sffamily}{}{}
% patch the patch by authblk so that the author block is flush left
\def\maketitle{{%
  \renewenvironment{tabular}[2][]
    {\begin{flushleft}}
    {\end{flushleft}}
  \AB@maketitle}}
\makeatletter
\renewcommand\AB@affilsepx{ \protect\Affilfont}
%\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{2pt}}
\renewcommand\AB@affilnote[1]{{\bfseries #1}\hspace{3pt}}
\renewcommand{\affil}[2][]%
   {\newaffiltrue\let\AB@blk@and\AB@pand
      \if\relax#1\relax\def\AB@note{\AB@thenote}\else\def\AB@note{#1}%
        \setcounter{Maxaffil}{0}\fi
        \begingroup
        \let\href=\href@Orig
        \let\texttt=\textttOrig
        \let\protect\@unexpandable@protect
        \def\thanks{\protect\thanks}\def\footnote{\protect\footnote}%
        \@temptokena=\expandafter{\AB@authors}%
        {\def\\{\protect\\\protect\Affilfont}\xdef\AB@temp{#2}}%
         \xdef\AB@authors{\the\@temptokena\AB@las\AB@au@str
         \protect\\[\affilsep]\protect\Affilfont\AB@temp}%
         \gdef\AB@las{}\gdef\AB@au@str{}%
        {\def\\{, \ignorespaces}\xdef\AB@temp{#2}}%
        \@temptokena=\expandafter{\AB@affillist}%
        \xdef\AB@affillist{\the\@temptokena \AB@affilsep
          \AB@affilnote{\AB@note}\protect\Affilfont\AB@temp}%
      \endgroup
       \let\AB@affilsep\AB@affilsepx
}
\makeatother
\renewcommand\Authfont{\sffamily\bfseries}
\renewcommand\Affilfont{\sffamily\small\mdseries}
\setlength{\affilsep}{1em}

\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{fontspec}

  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}

\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={gwsnr: A Python Package for Efficient SNR Calculation of Gravitational Waves},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% --- We redefined \texttt, but in sections and captions we want the
% --- old definition
\let\addcontentslineOrig=\addcontentsline
\def\addcontentsline#1#2#3{\bgroup
  \let\texttt=\textttOrig\addcontentslineOrig{#1}{#2}{#3}\egroup}
\let\markbothOrig\markboth
\def\markboth#1#2{\bgroup
  \let\texttt=\textttOrig\markbothOrig{#1}{#2}\egroup}
\let\markrightOrig\markright
\def\markright#1{\bgroup
  \let\texttt=\textttOrig\markrightOrig{#1}\egroup}

\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\title{\texttt{gwsnr}: A Python Package for Efficient SNR Calculation of
Gravitational Waves}

        \author[1]{Phurailatpam Hemantakumar}
          \author[1]{Otto Akseli HANNUKSELA}

      \affil[1]{Department of Physics, The Chinese University of Hong
Kong, Shatin, New Territories, Hong Kong}
  \date{\vspace{-7ex}}

\begin{document}
\maketitle

\marginpar{

  \begin{flushleft}
  %\hrule
  \sffamily\small

  {\bfseries DOI:} \href{https://doi.org/10.xxxxxx/draft}{\color{linky}{10.xxxxxx/draft}}

  \vspace{2mm}

  {\bfseries Software}
  \begin{itemize}
    \setlength\itemsep{0em}
    \item \href{}{\color{linky}{Review}} \ExternalLink
    \item \href{https://github.com/hemantaph/ler}{\color{linky}{Repository}} \ExternalLink
    \item \href{https://zenodo.org/badge/latestdoi/626733473}{\color{linky}{Archive}} \ExternalLink
  \end{itemize}

  \vspace{2mm}

  \par\noindent\hrulefill\par

  \vspace{2mm}

  {\bfseries Editor:} \href{}{Pending Editor} \ExternalLink \\
  \vspace{1mm}
    {\bfseries Reviewers:}
  \begin{itemize}
  \setlength\itemsep{0em}
    \item \href{https://github.com/Pending Reviewer}{@Pending Reviewer}
    \item \href{https://github.com/}{@}
    \end{itemize}
    \vspace{2mm}

  {\bfseries Submitted:} 3rd July 2024\\
  {\bfseries Published:} 3rd July 2024

  \vspace{2mm}
  {\bfseries License}\\
  Authors of papers retain copyright and release the work under a Creative Commons Attribution 4.0 International License (\href{http://creativecommons.org/licenses/by/4.0/}{\color{linky}{CC BY 4.0}}).

  \end{flushleft}
}

\section{Summary}\label{summary}

Gravitational waves, ripples in spacetime predicted by Einstein's theory
of General Relativity, have revolutionized astrophysics since their
first detection in 2015 (Abbott, B. P. et al. (2016a)). These waves are
emitted by cataclysmic events like the merging of binary black holes
(BBHs), binary neutron stars (BNSs) and BH-NS pairs, providing a unique
window into the cosmos. A critical aspect of gravitational wave analysis
is the Signal-to-Noise Ratio (SNR). SNR quantifies the strength of a
gravitational wave signal relative to the background noise in a
detector, like LIGO (The LIGO Scientific Collaboration et al. (2015)) or
Virgo (Acernese et al. (2014)). This ratio is pivotal in confirming the
detection of gravitational waves and extracting astrophysical
information from them (Abbott, B. P. et al. (2016b)). However, specific
scenarios in gravitational wave research, particularly in simulations of
detectable gravitational wave events (Abbott et al. (2016)) and in
hierarchical Bayesian analysis (Thrane and Talbot (2019)) where
selection effects are considered, demand extensive and efficient
computation of SNR. This requirement presents a significant challenge,
as conventional computational approaches are typically time-consuming
and impractical for such specialized and large-scale analyses (Taylor
and Gerosa (2018), Davide, Pratten, and Vecchio (2020)).

\section{Statement of Need}\label{statement-of-need}

The \emph{\texttt{gwsnr}} Python package addresses the need for
efficient SNR computation in gravitational wave research. It
innovatively streamlines SNR calculations, enhancing accuracy and
efficiency with several advanced techniques. Firstly, it utilizes an
innovative interpolation method, employing a partial-scaling approach
for accurately interpolating the SNR of gravitational waves from
spin-less binary systems wrt to their mass parameters. Secondly, the
package features a noise-weighted inner product method, similar to that
in the \emph{\texttt{bilby}} package (Ashton, Gregory and Biscoveanu,
Sylvia and Cornish, Neil and Dal Canto, Isaac and Kumar, Prayush and
Meacher, Duncan and Middleton, Hannah and Mistry, Divyansh and Smith,
Rory and Stevenson, Tom (2022)), but enhanced with multiprocessing
capabilities. This integration allows for the parallel processing of
complex calculations, thereby expediting the SNR computation. Thirdly, a
trained Artificial Neural Network (ANN) model is incorporated for rapid
`probability of detection' (Pdet) estimation for BBH systems with spin
precession. Lastly, \emph{\texttt{gwsnr}} leverages the
\emph{\texttt{numba}}'s Just-In-Time (njit) compiler (Lam, Pitrou, and
Seibert (2022)), which optimizes performance by compiling Python code
into machine code at runtime, drastically reducing execution times.
Beyond these technical merits, \emph{\texttt{gwsnr}} stands out for its
user-friendly features and seamless integration with other related
software packages, making it not just a powerful tool but also an
accessible one for researchers. These attributes position
\emph{\texttt{gwsnr}} as an invaluable asset in gravitational wave data
analysis, particularly in simulations of detectable binary mergers,
calculation of merger rates, determining gravitational wave lensing
rates ((\textbf{ler?}), Ng et al. (2024)), and in the analysis of
selection effects within hierarchical Bayesian frameworks. The package
thus represents a significant step forward in gravitational wave
research, enabling more precise and efficient exploration of the
universe through gravitational wave observations and simulations.
Additionally, \emph{\texttt{gwsnr}} is instrumental in the
\emph{\texttt{ler}} ((\textbf{ler?})) package for calculating detectable
rates of both lensed and unlensed gravitational wave events, showcasing
its utility in advanced gravitational wave studies.

\section{Mathematical Formulation}\label{mathematical-formulation}

\paragraph{Modified FINDCHIRP Method: Partial Scaling
Approach}\label{modified-findchirp-method-partial-scaling-approach}

The \emph{\texttt{gwsnr}} package introduces the Partial Scaling method
for SNR calculations in spin-less binary systems. This method, rooted in
the FINDCHIRP algorithm (Allen et al. (2012)), focuses on non-spinning
IMR waveforms and particularly interpolates the Partial scaled SNR
(\(\rho_{1/2}\)) based on mass parameters (\(M_{tot},q\)).

\begin{itemize}
\item
  \textbf{Interpolation Method}: Utilizes a 2D cubic spline technique
  (njit-ted) for the `partialsnr' segment.
\item
  \textbf{Equations}:

  \begin{itemize}
  \item
    For a simple inspiral waveform, the optimal SNR is given by,
    \[\rho = F(D_l,\mathcal{M},\iota,\psi,\alpha, \delta, \psi) \sqrt{ 4\int_0^{f_{ISCO}} \frac{f^{-7/3}}{S_n(f)}df }\]
  \item
    \(F\) is defined as a function of luminosity distance (\(D_l\)),
    chirp mass (\(\mathcal{M}\)), inclination angle (\(\iota\)),
    polarization angles (\(\psi\)), right ascension (\(\alpha\)), and
    declination (\(\delta\)). \(f\) is the frequency, \(f_{ISCO}\) is
    the last stable orbit frequency and \(S_n(f)\) is the detector's
    noise curve or power spectral density (psd).
  \item
    Then, partial scaled SNR:
    \(\rho_{1/2} = \sqrt{ 4\int_0^\infty \frac{f^{-7/3}}{S_n(f)}df } \approx \sqrt{ 4\int_0^{f_{ISCO}} \frac{f^{-7/3}}{S_n(f)}df }\)
  \item
    For an spinless frequency-domain IMR (Inspiral-Merger-Ringdown)
    waveform with optimal SNR equal to \(\rho\):
    \(\rho_{1/2} = \rho\,/\, F(D_l,\mathcal{M},\iota,\psi,\alpha, \delta, \psi)\)
  \item
    \(\rho_{1/2}\) is considered a function of \(M_{tot}\) and \(q\).
  \end{itemize}
\end{itemize}

\paragraph{Noise-Weighted Inner Product
Method}\label{noise-weighted-inner-product-method}

Designed for SNR calculations in systems characterized by frequency
domain waveforms in \emph{\texttt{lalsimulation}} (LIGO Scientific
Collaboration, Virgo Collaboration, and KAGRA Collaboration (2018)),
including spin-precessing binary systems.

\begin{itemize}
\item
  \textbf{Methodology}: Combines waveform generation (multi-process),
  antenna pattern calculation (njit-ted), and noise-weighted inner
  product computation (njit-ted).
\item
  \textbf{Equations}:

  \begin{itemize}
  \item
    Inner product:
    \(\left< a | b \right> = 4 \int_{f_{min}}^{f_{max}} \frac{\tilde{a}(f)\tilde{b}^*(f)}{S_n(f)} df\)
  \item
    Optimal SNR:
    \(\rho = \sqrt{ F_+^2 \left< \tilde{h}_+ | \tilde{h}_+ \right> + F_{\times}^2 \left< \tilde{h}_{\times} | \tilde{h}_{\times} \right> }\),
    for orthogonal \(h_+\) and \(h_{\times}\).
  \item
    \(h_{+\times}\) are frequency domain waveform polarizations, and
    \(F_{+\times}\) are antenna patterns.
  \end{itemize}
\end{itemize}

These formulations highlight \emph{\texttt{gwsnr}}'s capability to
efficiently process diverse gravitational wave signals, enhancing data
analysis accuracy and efficiency.

\paragraph{Artificial Neural Network (ANN) Model for Pdet
Estimation}\label{artificial-neural-network-ann-model-for-pdet-estimation}

The \emph{\texttt{gwsnr}} package now incorporates an artificial neural
network (ANN) model, developed using \emph{\texttt{TensorFlow}} (Abadi
et al. (2015)) and \emph{\texttt{sklearn}} (Pedregosa et al. (2011)), to
rapidly estimate the `probability of detection' (Pdet) in binary black
hole (BBH) systems using the IMRPhenomXPHM waveform approximant. This
complex inspiral-merger-ringdown (IMR) waveform model accounts for
spin-precessing systems with subdominant harmonics. The ANN model is
especially useful when precise signal-to-noise ratio (SNR) calculations
are not critical, providing a quick and effective means of estimating
Pdet. This value indicates detectability under Gaussian noise by
determining if the SNR exceeds a certain threshold. Trained on a large
dataset from the \emph{\texttt{ler}} package, the ANN model uses
`partial scaled SNR' values as a primary input, reducing input
dimensionality from 15 to 5 and enhancing accuracy. This approach offers
a practical solution for assessing detectability under specified
conditions.

\section{Acknowledgements}\label{acknowledgements}

The author gratefully acknowledges the substantial contributions from
all who supported this research. Special thanks go to my academic
advisors for their invaluable guidance and unwavering support. The
interactions with my research colleagues have greatly enriched this
work. The Department of Physics at The Chinese University of Hong Kong
is acknowledged for the Postgraduate Studentship that made this research
possible. Thanks are also due to the LIGO Laboratory for the
computational resources, supported by National Science Foundation Grants
No.~PHY-0757058 and No.~PHY-0823459.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-tensorflow2015}
Abadi, Martín, Ashish Agarwal, Paul Barham, Eugene Brevdo, Zhifeng Chen,
Craig Citro, Greg S. Corrado, et al. 2015. {``{TensorFlow}: Large-Scale
Machine Learning on Heterogeneous Systems.''}
\url{https://www.tensorflow.org/}.

\bibitem[\citeproctext]{ref-Abbott:2016:rates}
Abbott, B. P., R. Abbott, T. D. Abbott, M. R. Abernathy, F. Acernese, K.
Ackley, C. Adams, et al. 2016. {``ASTROPHYSICAL IMPLICATIONS OF THE
BINARY BLACK HOLE MERGER GW150914.''} \emph{The Astrophysical Journal
Letters} 818 (2): L22.
\url{https://doi.org/10.3847/2041-8205/818/2/l22}.

\bibitem[\citeproctext]{ref-Abbott:2016}
Abbott, B. P., R. Abbott, T. D. Abbott, M. R. Abernathy, F. Acernese, K.
Ackley, C. Adams, et al. 2016a. {``Observation of Gravitational Waves
from a Binary Black Hole Merger.''} \emph{Physical Review Letters} 116
(6). \url{https://doi.org/10.1103/physrevlett.116.061102}.

\bibitem[\citeproctext]{ref-Abbott:2016:detection}
Abbott, B. P., R. Abbott, T. D. Abbott, M. R. Abernathy, F. Acernese, K.
Ackley, C. Adams, et al. 2016b. {``GW150914: First Results from the
Search for Binary Black Hole Coalescence with Advanced LIGO.''}
\emph{Physical Review D} 93 (12).
\url{https://doi.org/10.1103/physrevd.93.122003}.

\bibitem[\citeproctext]{ref-VIRGO2015}
Acernese, F, M Agathos, K Agatsuma, D Aisa, N Allemandou, A Allocca, J
Amarni, et al. 2014. {``Advanced Virgo: A Second-Generation
Interferometric Gravitational Wave Detector.''} \emph{Classical and
Quantum Gravity} 32 (2): 024001.
\url{https://doi.org/10.1088/0264-9381/32/2/024001}.

\bibitem[\citeproctext]{ref-Allen:2012}
Allen, Bruce, Warren G. Anderson, Patrick R. Brady, Duncan A. Brown, and
Jolien D. E. Creighton. 2012. {``FINDCHIRP: An Algorithm for Detection
of Gravitational Waves from Inspiraling Compact Binaries.''}
\emph{Physical Review D} 85 (12).
\url{https://doi.org/10.1103/physrevd.85.122006}.

\bibitem[\citeproctext]{ref-bilby}
Ashton, Gregory and Biscoveanu, Sylvia and Cornish, Neil and Dal Canto,
Isaac and Kumar, Prayush and Meacher, Duncan and Middleton, Hannah and
Mistry, Divyansh and Smith, Rory and Stevenson, Tom. 2022. {``{bilby: a
user-friendly Bayesian inference library}.''} \emph{GitHub Repository}.
GitHub. \url{https://github.com/GregoryAshton/Bilby}.

\bibitem[\citeproctext]{ref-Gerosa:2020}
Davide, Gerosa, Geraint Pratten, and Alberto Vecchio. 2020.
{``Gravitational-Wave Selection Effects Using Neural-Network
Classifiers.''} \emph{Physical Review D} 102 (10).
\url{https://doi.org/10.1103/physrevd.102.103020}.

\bibitem[\citeproctext]{ref-numba}
Lam, Stan, Stéphane Pitrou, and Mark Seibert. 2022. {``Numba: A High
Performance Python Compiler.''} \emph{Numba Documentation}. Anaconda,
Inc. \url{https://numba.pydata.org/}.

\bibitem[\citeproctext]{ref-lalsuite}
LIGO Scientific Collaboration, Virgo Collaboration, and KAGRA
Collaboration. 2018. {``{LVK} {A}lgorithm {L}ibrary - {LALS}uite.''}
Free software (GPL). \url{https://doi.org/10.7935/GT1W-FZ16}.

\bibitem[\citeproctext]{ref-ng2024}
Ng, Leo C. Y., Justin Janquart, Hemantakumar Phurailatpam, Harsh Narola,
Jason S. C. Poon, Chris Van Den Broeck, and Otto A. Hannuksela. 2024.
{``Uncovering Faint Lensed Gravitational-Wave Signals and Reprioritizing
Their Follow-up Analysis Using Galaxy Lensing Forecasts with Detected
Counterparts.''} \url{https://arxiv.org/abs/2403.16532}.

\bibitem[\citeproctext]{ref-scikitlearn}
Pedregosa, Fabian, Gaël Varoquaux, Alexandre Gramfort, Vincent Michel,
Bertrand Thirion, Olivier Grisel, Mathieu Blondel, et al. 2011.
{``Scikit-Learn: Machine Learning in {P}ython.''} \emph{Journal of
Machine Learning Research} 12: 2825--30.

\bibitem[\citeproctext]{ref-Taylor:2018}
Taylor, Stephen R., and Davide Gerosa. 2018. {``Mining
Gravitational-Wave Catalogs to Understand Binary Stellar Evolution: A
New Hierarchical Bayesian Framework.''} \emph{Physical Review D} 98 (8).
\url{https://doi.org/10.1103/physrevd.98.083017}.

\bibitem[\citeproctext]{ref-LIGO2015}
The LIGO Scientific Collaboration, J Aasi, B P Abbott, R Abbott, T
Abbott, M R Abernathy, K Ackley, et al. 2015. {``Advanced LIGO.''}
\emph{Classical and Quantum Gravity} 32 (7): 074001.
\url{https://doi.org/10.1088/0264-9381/32/7/074001}.

\bibitem[\citeproctext]{ref-Thrane:2019}
Thrane, Eric, and Colm Talbot. 2019. {``An Introduction to Bayesian
Inference in Gravitational-Wave Astronomy: Parameter Estimation, Model
Selection, and Hierarchical Models.''} \emph{Publications of the
Astronomical Society of Australia} 36.
\url{https://doi.org/10.1017/pasa.2019.2}.

\end{CSLReferences}

\end{document}